---
title: "Preliminary Analysis"
output: html_document
date: "2025-12-04"
---
## Setups
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(foreign)
library(lavaan)
library(lme4)
library("dagitty")
library("semPower")
library("semPlot")
library(lubridate)
library(hms)
library("MplusAutomation")
library("BiocManager")
library("rhdf5")
EMA_final <- readRDS("EMA_final.rds")
glimpse(EMA_final)
```

## Sample size and specific sample sizes.
```{r}
#Counting final sample size
EMA_final %>%
  summarise(n = n_distinct(ID)) 
#Breakdown of group membership
EMA_final %>%
  group_by(Group) %>%
  summarise(n = n_distinct(ID))
#Prompts completed by each ppt
EMA_final %>%
  group_by(ID) %>%
  summarise(n_prompts = n(),
            compliance = n_prompts/70) %>%
  ungroup()


```

## Descriptives of SI
```{r}
#Proportion of prompts containing each type of SI
EMA_final %>%
  summarise(
    prop_any = mean(SI_any == 1, na.rm = TRUE),
    prop_pass = mean(passiveSI == 1, na.rm = TRUE),
    prop_act = mean(activeSI == 1, na.rm = TRUE),
    prop_SMI = mean(SMI == 1, na.rm = TRUE)
  )
```

## Descriptives of events
```{r}
#Proportion of prompts containing events
EMA_final %>%
  summarise(
    prop_stress = mean(stress_any == 1, na.rm = TRUE),
    prop_stressint1 = mean(stress_int == 1, na.rm = TRUE),
    prop_stressint2 = mean(stress_int == 2, na.rm = TRUE),
    prop_pos = mean(pos_any == 1, na.rm = TRUE),
    prop_posint1 = mean(pos_int == 1, na.rm = TRUE),
    prop_posint2 = mean(pos_int == 2, na.rm = TRUE)
  )
#Measuring collinearity of stress and positivity
#First making a table with the two events
tab_events <- with(EMA_final, table(stress_any, pos_any))
tab_events
prop.table(tab_events)           # proportions of all prompts
prop.table(tab_events, 1)        
prop.table(tab_events, 2)        
#Now using the table to measure how % stress so-occurred with pos and vice versa
n_both   <- tab_events["1", "1"]
n_total  <- sum(tab_events)
both <- n_both / n_total 

n_stress <- sum(tab_events["1", ])
stress_with_pos <- n_both / n_stress

n_pos    <- sum(tab_events[ , "1"])
pos_with_stress <- n_both / n_pos


```

## Descriptives of affect
```{r}
#glimpse at affect
summary(EMA_final$current_affect)
#affect breakdown by SI
EMA_final %>%
  group_by(SI_any) %>%
  summarise(
    mean_affect = mean(current_affect, na.rm = TRUE),
    sd_affect = sd(current_affect, na.rm = TRUE)
  )
#Okay, ppts had a general negative mood when SI
#Affects in stress - negative event-related affect
EMA_final %>%
  group_by(stress_int) %>%
  summarise(
    nNA = n(),
    mean_NA = mean(current_affect, na.rm = TRUE),
    sd_NA = sd(current_affect, na.rm = TRUE)
  ) %>%
  ungroup()
#Affect in pleasantness - positive event-related affect
EMA_final %>%
  group_by(pos_int) %>%
  summarise(
    nPA = n(),
    mean_PA = mean(current_affect, na.rm = TRUE),
    sd_PA = sd(current_affect, na.rm = TRUE)
  ) %>%
  ungroup()
#Hmmm...even though we have collinearity, but ppts still seem more happy in positive events and less happy in negative events.
```
## Check missingness
```{r}

sapply(EMA_final, function(x) mean(is.na(x)))

```

## Computing the (T-1) variables, disabling overnight lag.
```{r}
#Mark every first prompt of a day
EMA_mplus <- EMA_final %>%
  arrange(ID, survey_dtime) %>%
  group_by(ID) %>%
  mutate(
    day = as.Date(survey_dtime),
    new_d = is.na(lag(day)) | day != lag(day)
  ) %>%
  ungroup()
#Create a new variable for affect(T-1) and SI(T-1), which will be input into Mplus as predictors. Disabling every first prompt of the day shall avoid overnight lag.
EMA_mplus <- EMA_mplus %>%
  mutate(
    affect_T_1 = if_else(new_d, NA_real_, lag(current_affect)),
    SI_T_1 = if_else(new_d, NA_real_, lag(SI_any))
  )
```

## Computing post-event affect(T).
```{r}
#Filtering out post-event affect. These will be outcome variables for each event type.
EMA_mplus <- EMA_mplus %>%
  mutate(
    affect_post_stress = if_else(post_stress_affect == 1L,
                                 current_affect,
                                 NA_real_),
    affect_post_pos = if_else(post_pos_affect == 1L,
                              current_affect,
                              NA_real_)
  )
glimpse(EMA_mplus)
```

## Counting valid post-event affects we have - valid pairs for DSEM per ppt
```{r}
#Counting valid post-stress prompts
EMA_mplus %>%
  summarise(
  n_post_stress = sum(!is.na(post_stress_affect)))
EMA_mplus %>%
  group_by(ID) %>%
  count(!is.na(post_stress_affect))

#counting valid post-positive prompts
EMA_mplus %>%
  summarise(
    n_post_pos = sum(!is.na(post_pos_affect))
  )
EMA_mplus %>%
  group_by(ID) %>%
  count(!is.na(post_pos_affect))

```


## Making a tibble ready to be put into Mplus for analysis of Aim 1 

```{r}
EMA_DSEM1 <- EMA_mplus %>%
  arrange(ID, survey_dtime) %>%
  mutate(
    # Time in hours since first prompt for each ID
    time_hr = as.numeric(
      difftime(survey_dtime, first(survey_dtime), units = "hours"),
      units = "hours"
    ),
    .by = ID
  ) %>%
  #Calculating interactions
  mutate(slo = as.integer(stress_int == 1L),
         shi = as.integer(stress_int == 2L)) %>% 
  mutate(sloSI = slo * SI_T_1,
         shiSI = shi * SI_T_1) %>%
    select(
    ID,
    affect = current_affect,
    si = SI_any, slo, shi, sloSI, shiSI, time = time_hr)
#Calculating interactions (getting pass the NAs so mplus doesn't complain)
EMA_DSEM1 <- EMA_DSEM1 %>%
  mutate(
    sloSI = slo * si,
    shiSI = shi * si
  )
EMA_DSEM1 <- EMA_DSEM1 %>%
  group_by(ID) %>%
  mutate(
    si_b = mean(si, na.rm = TRUE),
    slo_b = mean(slo, na.rm = TRUE),
    shi_b = mean(shi, na.rm = TRUE),
    sloSI_b = mean(sloSI, na.rm = TRUE),
    shiSI_b = mean(shiSI, na.rm = TRUE)
  ) 

#Making data table
write.table(
  EMA_DSEM1,
  "EMA_DSEM1.dat",
  row.names = FALSE,
  col.names = FALSE,
  na = "-9999",
  quote = FALSE
)
glimpse(EMA_DSEM1)
```
## Now trying to make a copy for AIM2
```{r}
EMA_DSEM2 <- EMA_mplus %>%
  arrange(ID, survey_dtime) %>%
  mutate(
# Time in hours since first prompt for each ID
    time_hr = as.numeric(
      difftime(survey_dtime, first(survey_dtime), units = "hours"),
      units = "hours"
    ),
    .by = ID
  ) %>%
  mutate(plo = as.integer(pos_int == 1L),
         phi = as.integer(pos_int == 2L)) %>% 
  mutate(ploSI = plo * SI_T_1,
         phiSI = phi * SI_T_1) %>%
  select(
    ID,
    affect = current_affect,
    si = SI_any, plo, phi, ploSI, phiSI, time = time_hr)
#Calculating interactions
EMA_DSEM2 <- EMA_DSEM2 %>%
  mutate(
    ploSI = plo * si,
    phiSI = phi * si
  )
EMA_DSEM2 <- EMA_DSEM2 %>%
  group_by(ID) %>%
  mutate(
    si_b = mean(si, na.rm = TRUE),
    plo_b = mean(plo, na.rm = TRUE),
    phi_b = mean(phi, na.rm = TRUE),
    ploSI_b = mean(ploSI, na.rm = TRUE),
    phiSI_b = mean(phiSI, na.rm = TRUE)
  )
#Making data table
write.table(
  EMA_DSEM2,
  "EMA_DSEM2.dat",
  row.names = FALSE,
  col.names = FALSE,
  na = "-9999",
  quote = FALSE
)
glimpse(EMA_DSEM2)
```
## Building Aim3 dataset
```{r}
EMA_mplus <- EMA_mplus %>%
  group_by(ID) %>%
  mutate(
    pSI_T_1 = lag(passiveSI),
    aSI_T_1 = lag(activeSI),
    smi_T_1 = lag(SMI)
  ) %>%
  ungroup()
EMA_DSEM3 <- EMA_mplus %>%
  arrange(ID, survey_dtime) %>%
  mutate(
# Time in hours since first prompt for each ID
    time_hr = as.numeric(
      difftime(survey_dtime, first(survey_dtime), units = "hours"),
      units = "hours"
    ),
    .by = ID
  ) %>%
  mutate(
    pos_pSI = pos_any * pSI_T_1,
    str_pSI = stress_any * pSI_T_1,
    pos_aSI = pos_any * aSI_T_1,
    str_aSI = stress_any * aSI_T_1,
    pos_smi = pos_any * smi_T_1,
    str_smi = stress_any * smi_T_1
  ) %>% #interaction built
  select(
    ID, 
    affect = current_affect,
    psi = passiveSI,
    asi = activeSI,
    smi = SMI,
    time = time_hr,
    pos = pos_any,
    str = stress_any,
    stint = stress_int,
    point = pos_int,
    ppsi = pos_pSI,
    spsi = str_pSI,
    pasi = pos_aSI,
    sasi = str_aSI,
    psmi = pos_smi,
    ssmi = str_smi,
    pSI_T_1,
    aSI_T_1,
    smi_T_1
  ) # This is the master AIM3 tibble.

# Breakdown into three SIs
EMA_DSEM3_p <- EMA_DSEM3 %>%
  group_by(ID) %>%
  mutate(
    psi_b = mean(psi, na.rm = TRUE),
    pos_b = mean(pos, na.rm = TRUE),
    str_b = mean(str, na.rm = TRUE),
    posSI_b = mean(ppsi, na.rm = TRUE),
    strSI_b = mean(spsi, na.rm = TRUE)
  ) %>%
  ungroup() %>% #Creating between variable
  select(ID, psi_b, pos_b, str_b, posSI_b, strSI_b, affect, psi, ppsi, spsi, time, pos, str)
write.table(
  EMA_DSEM3_p,
  "EMA_DSEM3_p.dat",
  row.names = FALSE,
  col.names = FALSE,
  na = "-9999",
  quote = FALSE
)


EMA_DSEM3_a <- EMA_DSEM3 %>%
  group_by(ID) %>%
  mutate(
    asi_b = mean(asi, na.rm = TRUE),
    pos_b = mean(pos, na.rm = TRUE),
    str_b = mean(str, na.rm = TRUE),
    posSI_b = mean(pasi, na.rm = TRUE),
    strSI_b = mean(sasi, na.rm = TRUE)
  ) %>%
  ungroup() %>% #Creating between variable
  select(ID, asi_b, pos_b, str_b, posSI_b, strSI_b, affect, asi, time, pos, str, pasi, sasi)

write.table(
  EMA_DSEM3_a,
  "EMA_DSEM3_a.dat",
  row.names = FALSE,
  col.names = FALSE,
  na = "-9999",
  quote = FALSE
)

EMA_DSEM3_smi <- EMA_DSEM3 %>%
  group_by(ID) %>%
  mutate(
    smi_b = mean(smi, na.rm = TRUE),
    pos_b = mean(pos, na.rm = TRUE),
    str_b = mean(str, na.rm = TRUE),
    posSI_b = mean(psmi, na.rm = TRUE),
    strSI_b = mean(ssmi, na.rm = TRUE)
  ) %>%
  ungroup() %>% #Creating between variable
  select(ID, smi_b, pos_b, str_b, posSI_b, strSI_b, affect, smi, time, pos, str, psmi, ssmi)
write.table(
  EMA_DSEM3_smi,
  "EMA_DSEM3_smi.dat",
  row.names = FALSE,
  col.names = FALSE,
  na = "-9999",
  quote = FALSE
)
```

## Post-hoc starts here
```{r}
EMA_DSEM3_ps <- EMA_DSEM3 %>% # This shall be for passive
  mutate(slo = as.integer(stint == 1L),
         shi = as.integer(stint == 2L)) %>% 
  mutate(slopSI = slo * pSI_T_1,
         shipSI = shi * pSI_T_1) %>%
  group_by(ID) %>%
  mutate(
    psi_b = mean(psi, na.rm = TRUE),
    slo_b = mean(slo, na.rm = TRUE),
    shi_b = mean(shi, na.rm = TRUE),
    slopSI_b = mean(slopSI, na.rm = TRUE),
    shipSI_b = mean(shipSI, na.rm = TRUE)
  ) %>%
  select(ID, affect, psi, slo, shi, slopSI, shipSI, time, psi_b, slo_b, shi_b, slopSI_b, shipSI_b)

write.table(
  EMA_DSEM3_ps,
  "EMA_DSEM3_ps.dat",
  row.names = FALSE,
  col.names = FALSE,
  na = "-9999",
  quote = FALSE
)
 
EMA_DSEM3_pp <- EMA_DSEM3 %>% # This shall be for passive
  mutate(plo = as.integer(point == 1L),
         phi = as.integer(point == 2L)) %>% 
  mutate(plopSI = plo * pSI_T_1,
         phipSI = phi * pSI_T_1) %>%
  group_by(ID) %>%
  mutate(
    psi_b = mean(psi, na.rm = TRUE),
    plo_b = mean(plo, na.rm = TRUE),
    phi_b = mean(phi, na.rm = TRUE),
    plopSI_b = mean(plopSI, na.rm = TRUE),
    phipSI_b = mean(phipSI, na.rm = TRUE)
  ) %>%
  select(ID, affect, psi, plo, phi, plopSI, phipSI, time, psi_b, plo_b, phi_b, plopSI_b, phipSI_b)

write.table(
  EMA_DSEM3_pp,
  "EMA_DSEM3_pp.dat",
  row.names = FALSE,
  col.names = FALSE,
  na = "-9999",
  quote = FALSE
)

EMA_DSEM3_as <- EMA_DSEM3 %>% # This shall be for passive
  mutate(slo = as.integer(stint == 1L),
         shi = as.integer(stint == 2L)) %>% 
  mutate(sloaSI = slo * aSI_T_1,
         shiaSI = shi * aSI_T_1) %>%
  group_by(ID) %>%
  mutate(
    asi_b = mean(asi, na.rm = TRUE),
    slo_b = mean(slo, na.rm = TRUE),
    shi_b = mean(shi, na.rm = TRUE),
    sloaSI_b = mean(sloaSI, na.rm = TRUE),
    shiaSI_b = mean(shiaSI, na.rm = TRUE)
  ) %>%
  select(ID, affect, asi, slo, shi, sloaSI, shiaSI, time, asi_b, slo_b, shi_b, sloaSI_b, shiaSI_b)

write.table(
  EMA_DSEM3_as,
  "EMA_DSEM3_as.dat",
  row.names = FALSE,
  col.names = FALSE,
  na = "-9999",
  quote = FALSE
)

EMA_DSEM3_ap <- EMA_DSEM3 %>% # This shall be for passive
  mutate(plo = as.integer(point == 1L),
         phi = as.integer(point == 2L)) %>% 
  mutate(ploaSI = plo * aSI_T_1,
         phiaSI = phi * aSI_T_1) %>%
  group_by(ID) %>%
  mutate(
    asi_b = mean(asi, na.rm = TRUE),
    plo_b = mean(plo, na.rm = TRUE),
    phi_b = mean(phi, na.rm = TRUE),
    ploaSI_b = mean(ploaSI, na.rm = TRUE),
    phiaSI_b = mean(phiaSI, na.rm = TRUE)
  ) %>%
  select(ID, affect, asi, plo, phi, ploaSI, phiaSI, time, asi_b, plo_b, phi_b, ploaSI_b, phiaSI_b)

write.table(
  EMA_DSEM3_ap,
  "EMA_DSEM3_ap.dat",
  row.names = FALSE,
  col.names = FALSE,
  na = "-9999",
  quote = FALSE
)

EMA_DSEM3_smis <- EMA_DSEM3 %>% # This shall be for passive
  mutate(slo = as.integer(stint == 1L),
         shi = as.integer(stint == 2L)) %>% 
  mutate(sloSI = slo * smi_T_1,
         shiSI = shi * smi_T_1) %>%
  group_by(ID) %>%
  mutate(
    smi_b = mean(smi, na.rm = TRUE),
    slo_b = mean(slo, na.rm = TRUE),
    shi_b = mean(shi, na.rm = TRUE),
    slosmi_b = mean(sloSI, na.rm = TRUE),
    shismi_b = mean(shiSI, na.rm = TRUE)
  ) %>%
  select(ID, affect, smi, slo, shi, sloSI, shiSI, time, smi_b, slo_b, shi_b, slosmi_b, shismi_b)

write.table(
  EMA_DSEM3_smis,
  "EMA_DSEM3_smis.dat",
  row.names = FALSE,
  col.names = FALSE,
  na = "-9999",
  quote = FALSE
)

EMA_DSEM3_smip <- EMA_DSEM3 %>% # This shall be for passive
  mutate(plo = as.integer(point == 1L),
         phi = as.integer(point == 2L)) %>% 
  mutate(plosmi = plo * smi_T_1,
         phismi = phi * smi_T_1) %>%
  group_by(ID) %>%
  mutate(
    smi_b = mean(smi, na.rm = TRUE),
    plo_b = mean(plo, na.rm = TRUE),
    phi_b = mean(phi, na.rm = TRUE),
    plosmi_b = mean(plosmi, na.rm = TRUE),
    phismi_b = mean(phismi, na.rm = TRUE)
  ) %>%
  select(ID, affect, smi, plo, phi, plosmi, phismi, time, smi_b, plo_b, phi_b, plosmi_b, phismi_b)

write.table(
  EMA_DSEM3_smip,
  "EMA_DSEM3_smip.dat",
  row.names = FALSE,
  col.names = FALSE,
  na = "-9999",
  quote = FALSE
)
```


```{r}
Baseline <- read.csv("DEETs2data_Baseline_cleaned 082425.csv")
Baseline %>%
  summarise(m_age = mean(BL_C_Age),
            sd_age = sd(BL_C_Age))
```


```{r}
df_interval <- EMA_final %>%
  mutate(
    survey_dtime = ymd_hms(survey_dtime, tz = "America/New_York")
  ) %>%
  arrange(ID, survey_dtime) %>%
  group_by(ID) %>%
  mutate(
    previous_prompt_time = lag(survey_dtime),
    interval_hours = as.numeric(difftime(
      survey_dtime,
      previous_prompt_time,
      units = "hours"
    )),
    prompt_date = as.Date(survey_dtime),
    previous_prompt_date = as.Date(previous_prompt_time),
    same_day_interval = prompt_date == previous_prompt_date
  ) %>%
  ungroup()

overall_interval_descriptives <- df_interval %>%
  summarise(
    n_intervals = sum(!is.na(interval_hours)),
    mean_interval_hours = mean(interval_hours, na.rm = TRUE),
    sd_interval_hours = sd(interval_hours, na.rm = TRUE),
    median_interval_hours = median(interval_hours, na.rm = TRUE),
    min_interval_hours = min(interval_hours, na.rm = TRUE),
    max_interval_hours = max(interval_hours, na.rm = TRUE)
  )

overall_interval_descriptives
```

